---
title: "GRH Forecast Evaluation"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: sandstone
    collapsed: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
# All evaluation points present in the data.
all_locs <- c("wilder", "bellows", "vernon", "ottauquechee", "sugar", "white")
# Only these locations are included in the figures and tables. The three
# tributaries are retained in the data because they contribute to the mainstem
# inflows, but the three dams are what the evaluation reports on.
locs <- c("wilder", "bellows", "vernon")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 3
)

library(tidyverse)
suppressMessages(import::from(
  hydroGOF,
  gof,
  NSE,
  KGE,
  me,
  mse,
  pbias,
  nrmse,
  rPearson,
  rSD,
  mNSE
))
import::from(kableExtra, kbl, kable_minimal, kable_styling)
```

Report created: `r now()`

## Metric Descriptions

- *NSE* -  Nash-Sutcliffe Efficiency, a metric that measures performance relative to the mean of the data, values range from [-∞ , 1], with values greater than 0 indicating better performance than the mean, and a perfect forecast will have a value of 1. 
- *KGE* - Kling-Gupta Efficiency, a metric that is often used similarly to NSE, but addresses several shortcomings by explicitly incorporating variability and correlation between simulations and observations. Values can range from [-∞ , 1] with negative values generally considered to indicate poor performance. 
- *PBIAS* - Percent Bias, Bias of the simulation relative to the observations, represented as a percent. Negative values indicate a low bias and positive values indicate a high bias, values closer to 0 are preferable. 
- *nRMSE* - Normalized root mean square error, expressed as a percent, lower values are better. 
- *r* - Pearson's correlation, a measure of how well the forecast and observations are correlated, values closer to 1 indicate better correlation. 
- *rSD* - Ratio of standard deviations of forecast and observed. Values close to 1 are preferable as they indicate that the forecast has the same variability as the observations. 


```{r, include=FALSE}
forecasts_0h <- read_csv(
  "processed-data/forecasts_00h.csv",
  show = F,
  progress = F
)
forecasts_6h <- read_csv(
  "processed-data/forecasts_06h.csv",
  show = F,
  progress = F
)
forecasts_12h <- read_csv(
  "processed-data/forecasts_12h.csv",
  show = F,
  progress = F
)
forecasts_24h <- read_csv(
  "processed-data/forecasts_24h.csv",
  show = F,
  progress = F
)
# added by Y.Son
target <- read_csv("processed-data/targets_24h.csv", show = F, progress = F) |>
  add_column(
    source = "GRH"
  ) |>
  rename(
    wilder_fc = wilder_tgt,
    bellows_fc = bellows_tgt,
    vernon_fc = vernon_tgt,
    white_fc = white_tgt,
    ottauquechee_fc = ottauquechee_tgt,
    sugar_fc = sugar_tgt
  )
forecasts_24h <- forecasts_24h |>
  bind_rows(target)

now <- now()

forecasts <- forecasts_24h |>
  mutate(
    source = case_when(
      source == "A" ~ "A",
      source == "B" ~ "B",
      source == "GRH" ~ "GRH",
      .default = NA
    )
  ) |>
  filter(
    issue_time >= ISOdate(2023, 6, 1, tz = "US/Eastern"),
    issue_time <=
      ISOdate(year(now), month(floor_date(now, "month")), 1, tz = "US/Eastern"),
  ) |>
  mutate(
    season = case_when(
      month(issue_time) %in% c(12, 1, 2) ~ "winter",
      month(issue_time) %in% 3:5 ~ "spring",
      month(issue_time) %in% 6:8 ~ "summer",
      month(issue_time) %in% 9:11 ~ "fall"
    )
  )

cols <- c("#9FBB73", "#F3B664", "#c44900", "#2C4E80", "#4793AF")

hourly <-
  forecasts |>
  group_by(issue_time) |>
  pivot_longer(
    -c(valid_time, issue_time, source, season),
    names_sep = "_",
    names_to = c("location", "type")
  ) |>
  mutate(location = factor(location, levels = all_locs)) |>
  mutate(
    lead_time = difftime(valid_time, issue_time, units = "hours") |>
      as.numeric()
  ) |>
  pivot_wider(
    id_cols = c(valid_time, issue_time, source, location, lead_time, season),
    names_from = type
  ) |>
  na.omit() |>
  mutate(error = fc - obs) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      mutate(
        source = "perfect",
        fc = obs,
        error = fc - obs
      )
  ) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      group_by(location, issue_time) |>
      mutate(
        source = "persistence",
        fc = obsprev[1],
        error = fc - obs
      )
  ) |>
  filter(location %in% locs)

forecasts |>
  distinct(valid_time, source, .keep_all = T) |>
  select(-issue_time, -season) |>
  pivot_longer(
    -c(valid_time, source),
    names_sep = "_",
    names_to = c("location", "type")
  ) |>
  mutate(valid_date = floor_date(valid_time, "day")) |>
  select(-valid_time) |>
  group_by(valid_date, location, type) |>
  mutate(
    location = factor(location, levels = all_locs)
  ) |>
  group_by(valid_date, source, location, type) |>
  summarise(value = mean(value), .groups = "drop") |>
  pivot_wider(id_cols = c(valid_date, source, location), names_from = type) |>
  mutate(obsprev = lag(obs)) |>
  select(-fc, -obs) |>
  na.omit() -> obs_prev_day

daily <-
  forecasts |>
  group_by(issue_time) |>
  mutate(valid_date = floor_date(valid_time, "day")) |>
  pivot_longer(
    -c(valid_time, issue_time, source, valid_date, season),
    names_sep = "_",
    names_to = c("location", "type")
  ) |>
  mutate(
    location = factor(location, levels = all_locs)
  ) |>
  group_by(valid_date, issue_time, source, location, type, season) |>
  summarise(value = mean(value), .groups = "drop") |>
  mutate(
    lead_time = difftime(valid_date, issue_time, units = "hours") |>
      as.numeric()
  ) |>
  left_join(obs_prev_day, by = join_by(valid_date, source, location)) |>
  mutate(value = ifelse(type == "obsprev", obsprev, value)) |>
  select(-obsprev) |>
  pivot_wider(
    id_cols = c(valid_date, issue_time, source, location, lead_time, season),
    names_from = type
  ) |>
  na.omit() |>
  mutate(error = fc - obs) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      mutate(
        source = "perfect",
        fc = obs,
        error = fc - obs
      )
  ) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      group_by(location, issue_time) |>
      mutate(
        source = "persistence",
        fc = obsprev[1],
        error = fc - obs
      )
  ) |>
  filter(location %in% locs)


n_day_ahead <- function(x, n) {
  x |> filter(lead_time <= n * 24, lead_time > (n - 1) * 24)
}

# revised by Y.Son
#day1_ave_forecasts <- daily |> n_day_ahead(1)
#day2_ave_forecasts <- daily |> n_day_ahead(2)
#day3_ave_forecasts <- daily |> n_day_ahead(3)
daily_offset_grh <- daily |>
  mutate(lead_time = ifelse(source == 'GRH', lead_time + 23, lead_time))
day1_ave_forecasts <- daily_offset_grh |> n_day_ahead(1)
day2_ave_forecasts <- daily_offset_grh |> n_day_ahead(2)
day3_ave_forecasts <- daily_offset_grh |> n_day_ahead(3)

fc_stats <- function(x) {
  x |>
    group_by(source, location, lead_time) |>
    mutate(n = n()) |>
    filter(n > 1) |>
    summarise(
      NSE = NSE(fc, obs),
      KGE = KGE(fc, obs),
      PBIAS = pbias(fc, obs),
      nRMSE = nrmse(fc, obs),
      r = rPearson(fc, obs),
      rSD = rSD(fc, obs),
      .groups = "drop"
    )
}

fc_stats_by_subgroup <- function(x, subgroup) {
  x |>
    mutate(month = month(issue_time)) |>
    group_by(source, location, lead_time, !!as.name(subgroup)) |>
    mutate(n = n()) |>
    filter(n > 1) |>
    summarise(
      NSE = NSE(fc, obs),
      KGE = KGE(fc, obs),
      PBIAS = pbias(fc, obs),
      nRMSE = nrmse(fc, obs),
      r = rPearson(fc, obs),
      rSD = rSD(fc, obs),
      .groups = "drop"
    )
}

fc_stats_daily <- daily |> fc_stats()
fc_stats_hourly <- hourly |> fc_stats()
fc_stats_da <- day1_ave_forecasts |>
  fc_stats() |>
  filter(lead_time >= 20)

fc_stats_da_season <- day1_ave_forecasts |>
  fc_stats_by_subgroup("season") |>
  filter(lead_time >= 20)

fc_stats_daily_by_month <- daily |> fc_stats_by_subgroup("month")
fc_stats_hourly_by_month <- hourly |> fc_stats_by_subgroup("month")
fc_stats_da_by_month <- day1_ave_forecasts |>
  fc_stats_by_subgroup("month") |>
  filter(lead_time >= 20)

fc_stats_daily_by_season <- daily |> fc_stats_by_subgroup("season")
fc_stats_hourly_by_season <- hourly |> fc_stats_by_subgroup("season")
```

### Day Ahead Forecast Statistics {.tabset}
#### Overall
```{r}
fc_stats_da |>
  arrange(location, source) |>
  select(-lead_time) |>
  # filter(!(source %in% c('perfect','persistence'))) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Winter 
```{r}
fc_stats_da_season |>
  filter(season == "winter") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Spring 
```{r}
fc_stats_da_season |>
  filter(season == "spring") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Summer 
```{r}
fc_stats_da_season |>
  filter(season == "summer") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Fall 
```{r}
fc_stats_da_season |>
  filter(season == "fall") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

### Error Distributions (Forecast - Observed) {.tabset}
```{r}
plot_error_boxplots <- function(fc) {
  fc |>
    group_by(source, location) |>
    filter(error < quantile(error, 0.95), error > quantile(error, 0.05)) |>
    bind_rows(fc |> filter(source == "perfect")) |>
    ggplot() +
    geom_hline(yintercept = 0) +
    geom_boxplot(aes(source, error, fill = source)) +
    facet_wrap(~location, scales = "free") +
    scale_fill_manual(values = cols) +
    theme_bw() +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 20, hjust = 1)
    ) +
    labs(y = "Daily Average Flow Error [cfs]", x = "Forecast Source")
}
```

#### Overall 
```{r}
day1_ave_forecasts |> plot_error_boxplots()
```

#### Winter 
```{r}
day1_ave_forecasts |>
  filter(season == "winter") |>
  plot_error_boxplots()
```

#### Spring 
```{r}
day1_ave_forecasts |>
  filter(season == "spring") |>
  plot_error_boxplots()
```

#### Summer 
```{r}
day1_ave_forecasts |>
  filter(season == "summer") |>
  plot_error_boxplots()
```

#### Fall 
```{r}
day1_ave_forecasts |>
  filter(season == "fall") |>
  plot_error_boxplots()
```


### Day Ahead Observed vs. Forecast Scatterplot {.tabset}
```{r}
plot_obs_vs_fc <- function(fc) {
  fc |>
    ggplot() +
    geom_point(aes(obs, fc, color = source)) +
    geom_abline(slope = 1) +
    theme_bw() +
    facet_wrap(~source) +
    scale_color_manual(values = cols) +
    theme(legend.position = "none") +
    coord_equal(xlim = c(0, 80000), ylim = c(0, 80000)) +
    labs(y = "Forecast", x = "Observed")
}
```

#### Overall 
```{r}
day1_ave_forecasts |> plot_obs_vs_fc()
```

#### Winter 
```{r}
day1_ave_forecasts |>
  filter(season == "winter") |>
  plot_obs_vs_fc()
```

#### Spring 
```{r}
day1_ave_forecasts |>
  filter(season == "spring") |>
  plot_obs_vs_fc()
```

#### Summer 
```{r}
day1_ave_forecasts |>
  filter(season == "summer") |>
  plot_obs_vs_fc()
```

#### Fall 
```{r}
day1_ave_forecasts |>
  filter(season == "fall") |>
  plot_obs_vs_fc()
```

### Day Ahead Performance by Month {.tabset}


```{r}
stat_by_month <- function(x, stat) {
  x |>
    mutate(plot_month = ISOdate(ifelse(month >= 6, 2000, 2001), month, 1)) |>
    ggplot() +
    # The no-skill threshold differs by metric: using the mean of the
    # observations as a predictor gives NSE = 0 but KGE = 1 - sqrt(2) ~ -0.41
    # (Knoben et al., 2019), so the reference line must be metric-specific.
    geom_hline(
      yintercept = ifelse(stat == "KGE", -0.41, 0),
      linetype = "dashed"
    ) +
    geom_line(aes(plot_month, !!as.name(stat), color = source), size = 1) +
    facet_wrap(~location) +
    scale_x_datetime(
      date_breaks = "1 month",
      date_labels = "%b",
      minor_breaks = NULL
    ) +
    scale_color_manual("Forecast Source", values = cols) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90)) +
    labs(x = "") +
    theme(legend.position = "top")
}
```

#### Kling-Gupta Efficiency (KGE)
```{r}
fc_stats_da_by_month |>
  stat_by_month("KGE") +
  coord_cartesian(ylim = c(-0.41, 1))
```

#### Nash-Suttecliffe Efficiency (NSE)
```{r}
fc_stats_da_by_month |>
  stat_by_month("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Pearson's Correlation (r)
```{r}
fc_stats_da_by_month |>
  stat_by_month("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Percent Bias
```{r}
fc_stats_da_by_month |>
  stat_by_month("PBIAS") +
  coord_cartesian(ylim = c(-100, 100))
```

## Daily forecast performance vs lead time {.tabset}

```{r}
stat_plot_over_time <- function(x, stat) {
  x |>
    ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_line(aes(lead_time / 24, !!as.name(stat), color = source), size = 1) +
    scale_x_continuous(breaks = 0:10, minor_breaks = NULL) +
    facet_wrap(~location) +
    theme_bw() +
    scale_color_manual("Forecast Source", values = cols) +
    # geom_vline(xintercept = 24) +
    theme(legend.position = "top") +
    labs(x = "Lead Time (days)")
}
```

### Kling-Gupta Efficiency (KGE) {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-0.41, 1))
```

#### Winter
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Spring
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Summer
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Fall
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

### Nash-Suttecliffe Efficiency (NSE) {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Winter
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Spring
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Summer
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Fall
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```


### Pearson's Correlation (r) {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Winter
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Spring
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Summer
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Fall
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```


### Percent Bias {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Winter 
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Spring 
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Summer 
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Fall 
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

## Hourly forecast performance vs lead time {.tabset}

### Kling-Gupta Efficiency (KGE) {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Winter
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Spring
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Summer
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Fall
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```
### Nash-Suttecliffe Efficiency (NSE) {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Winter
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Spring
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Summer
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Fall
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

### Pearson's Correlation (r) {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Winter
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Spring
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Summer
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Fall
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```


### Percent Bias {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Winter 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Spring 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Summer 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Fall 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

## Cumulative Error {.tabset}
```{r}
threshold <- 15000
threshold_elev <- 5 / 10
cum_error_raw <- hourly |>
  filter(lead_time > 0) |>
  arrange(location, source, issue_time, lead_time) |>
  group_by(location, source, issue_time) |>
  mutate(
    cum_error = cumsum(error),
    dist_to_upper_thresh = pmin(pmax(threshold - cum_error, 0), 2 * threshold),
    dist_to_lower_thresh = pmin(pmax(cum_error + threshold, 0), 2 * threshold),
    dist_to_upper_thresh = (dist_to_upper_thresh - threshold) /
      dist_to_upper_thresh *
      5 /
      10,
    dist_to_lower_thresh = (dist_to_lower_thresh - threshold) /
      dist_to_lower_thresh *
      5 /
      10,
    dist_to_nearest_thresh = pmin(dist_to_upper_thresh, dist_to_lower_thresh)
  ) |>
  ungroup()

cum_error_raw_seasonal <- hourly |>
  filter(lead_time > 0) |>
  arrange(location, source, issue_time, lead_time) |>
  group_by(location, source, issue_time, season) |>
  mutate(
    cum_error = cumsum(error),
    dist_to_upper_thresh = pmin(pmax(threshold - cum_error, 0), 2 * threshold),
    dist_to_lower_thresh = pmin(pmax(cum_error + threshold, 0), 2 * threshold),
    dist_to_upper_thresh = (dist_to_upper_thresh - threshold) /
      dist_to_upper_thresh *
      5 /
      10,
    dist_to_lower_thresh = (dist_to_lower_thresh - threshold) /
      dist_to_lower_thresh *
      5 /
      10,
    dist_to_nearest_thresh = pmin(dist_to_upper_thresh, dist_to_lower_thresh)
  ) |>
  ungroup()

cum_error_raw |>
  group_by(location, lead_time, source) |>
  summarise(
    ave_cum_error = mean(cum_error),
    q95_cum_error = quantile(cum_error, 0.975),
    q75_cum_error = quantile(cum_error, 0.75),
    q50_cum_error = quantile(cum_error, 0.50),
    q25_cum_error = quantile(cum_error, 0.25),
    q05_cum_error = quantile(cum_error, 0.025),
    p_bust = length(which(cum_error > threshold | cum_error < -threshold)) /
      length(cum_error),
    dist_to_upper_thresh_q95 = quantile(dist_to_upper_thresh, .975),
    dist_to_upper_thresh_q75 = quantile(dist_to_upper_thresh, .75),
    dist_to_upper_thresh_q50 = quantile(dist_to_upper_thresh, .5),
    dist_to_upper_thresh_q25 = quantile(dist_to_upper_thresh, .25),
    dist_to_upper_thresh_q05 = quantile(dist_to_upper_thresh, .025),
    dist_to_lower_thresh_q95 = quantile(dist_to_lower_thresh, .975),
    dist_to_lower_thresh_q75 = quantile(dist_to_lower_thresh, .75),
    dist_to_lower_thresh_q50 = quantile(dist_to_lower_thresh, .5),
    dist_to_lower_thresh_q25 = quantile(dist_to_lower_thresh, .25),
    dist_to_lower_thresh_q05 = quantile(dist_to_lower_thresh, .025),
    dist_to_nearest_thresh_q95 = quantile(dist_to_nearest_thresh, .975),
    dist_to_nearest_thresh_q75 = quantile(dist_to_nearest_thresh, .75),
    dist_to_nearest_thresh_q50 = quantile(dist_to_nearest_thresh, .5),
    dist_to_nearest_thresh_q25 = quantile(dist_to_nearest_thresh, .25),
    dist_to_nearest_thresh_q05 = quantile(dist_to_nearest_thresh, .025),
    .groups = "drop"
  ) |>
  filter(location %in% c("wilder", "vernon", "bellows")) -> cum_error

cum_error_raw_seasonal |>
  group_by(location, lead_time, source, season) |>
  summarise(
    ave_cum_error = mean(cum_error),
    q95_cum_error = quantile(cum_error, 0.975),
    q75_cum_error = quantile(cum_error, 0.75),
    q50_cum_error = quantile(cum_error, 0.50),
    q25_cum_error = quantile(cum_error, 0.25),
    q05_cum_error = quantile(cum_error, 0.025),
    p_bust = length(which(cum_error > threshold | cum_error < -threshold)) /
      length(cum_error),
    dist_to_upper_thresh_q95 = quantile(dist_to_upper_thresh, .975),
    dist_to_upper_thresh_q75 = quantile(dist_to_upper_thresh, .75),
    dist_to_upper_thresh_q50 = quantile(dist_to_upper_thresh, .5),
    dist_to_upper_thresh_q25 = quantile(dist_to_upper_thresh, .25),
    dist_to_upper_thresh_q05 = quantile(dist_to_upper_thresh, .025),
    dist_to_lower_thresh_q95 = quantile(dist_to_lower_thresh, .975),
    dist_to_lower_thresh_q75 = quantile(dist_to_lower_thresh, .75),
    dist_to_lower_thresh_q50 = quantile(dist_to_lower_thresh, .5),
    dist_to_lower_thresh_q25 = quantile(dist_to_lower_thresh, .25),
    dist_to_lower_thresh_q05 = quantile(dist_to_lower_thresh, .025),
    dist_to_nearest_thresh_q95 = quantile(dist_to_nearest_thresh, .975),
    dist_to_nearest_thresh_q75 = quantile(dist_to_nearest_thresh, .75),
    dist_to_nearest_thresh_q50 = quantile(dist_to_nearest_thresh, .5),
    dist_to_nearest_thresh_q25 = quantile(dist_to_nearest_thresh, .25),
    dist_to_nearest_thresh_q05 = quantile(dist_to_nearest_thresh, .025),
    .groups = "drop"
  ) |>
  filter(location %in% c("wilder", "vernon", "bellows")) -> cum_error_seasonal
```

### 2.5th to 97.5th percentile (Middle 95%) {.tabset}
```{r}
cum_error_plot <- function(error, threshold = threshold_elev, scale = 30000) {
  ggplot(error) +
    geom_hline(yintercept = c(-threshold, threshold), linetype = "dashed") +
    geom_errorbar(aes(
      lead_time / 24,
      y = q50_cum_error / scale,
      ymin = q05_cum_error / scale,
      ymax = q95_cum_error / scale,
      color = source
    )) +
    geom_point(aes(lead_time / 24, y = q50_cum_error / scale, color = source)) +
    scale_color_manual("Forecast Source", values = cols) +
    scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
    scale_y_continuous(
      breaks = seq(-threshold * 1.2, threshold * 1.2, len = 7)
    ) +
    coord_cartesian(ylim = c(-threshold * 1.2, threshold * 1.2)) +
    facet_grid(source ~ location) +
    theme_bw() +
    theme(
      legend.position = "top",
      panel.grid.minor = element_blank()
    ) +
    labs(
      x = "Lead Time (days)",
      y = "Cumulative Error (ft)"
    )
}
```

#### Overall
```{r, fig.height=8}
cum_error |>
  cum_error_plot()
```


#### Winter
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "winter") |>
  cum_error_plot()
```

#### Spring
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "spring") |>
  cum_error_plot()
```

#### Summer
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "summer") |>
  cum_error_plot()
```

#### Fall
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "fall") |>
  cum_error_plot()
```

### 25th to 75th percentile (Middle 50%) {.tabset}
```{r}
cum_error_plot50 <- function(error, threshold = threshold_elev, scale = 30000) {
  ggplot(error) +
    geom_hline(yintercept = c(-threshold, threshold), linetype = "dashed") +
    geom_errorbar(aes(
      lead_time / 24,
      y = q50_cum_error / scale,
      ymin = q25_cum_error / scale,
      ymax = q75_cum_error / scale,
      color = source
    )) +
    geom_point(aes(lead_time / 24, y = q50_cum_error / scale, color = source)) +
    scale_color_manual("Forecast Source", values = cols) +
    scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
    scale_y_continuous(
      breaks = seq(-threshold * 1.2, threshold * 1.2, len = 7)
    ) +
    coord_cartesian(ylim = c(-threshold * 1.2, threshold * 1.2)) +
    facet_grid(source ~ location) +
    theme_bw() +
    theme(
      legend.position = "top",
      panel.grid.minor = element_blank()
    ) +
    labs(
      x = "Lead Time (days)",
      y = "Cumulative Error (ft)"
    )
}
```

#### Overall
```{r, fig.height=8}
cum_error |>
  cum_error_plot50()
```


#### Winter
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "winter") |>
  cum_error_plot50()
```

#### Spring
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "spring") |>
  cum_error_plot50()
```

#### Summer
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "summer") |>
  cum_error_plot50()
```

#### Fall
```{r, fig.height=8}
cum_error_seasonal |>
  filter(season == "fall") |>
  cum_error_plot50()
```

```{r}
### 25th to 75th percentile (Middle 50%)
# ggplot(cum_error) +
#   geom_hline(yintercept = c(-threshold, threshold), linetype = "dashed") +
#   # geom_hline(yintercept = c(threshold, -2*threshold), linetype = "dashed", color = "red") +
#   geom_errorbar(aes(lead_time / 24, y = q50_cum_error, ymin = q25_cum_error, ymax = q75_cum_error, color = source)) +
#   geom_point(aes(lead_time / 24, y = q50_cum_error, color = source)) +
#   scale_color_manual("Forecast Source", values = cols) +
#   scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
#   coord_cartesian(ylim = c(-threshold, threshold)) +
#   facet_grid(source ~ location) +
#   theme_bw() +
#   theme(legend.position = "top") +
#   labs(
#     x = "Lead Time (days)",
#     y = "Cumulative Error (cfs)"
#   )
```

## Probability of exceeding 0.5 ft cumulative error {.tabset}
```{r}
prob_bust_plot <- function(error) {
  ggplot(error) +
    geom_line(aes(lead_time / 24, y = p_bust, color = source), size = 1) +
    scale_color_manual("Forecast Source", values = cols) +
    scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
    scale_y_continuous(labels = scales::label_percent()) +
    facet_grid(~location) +
    theme_bw() +
    theme(legend.position = "top") +
    labs(
      x = "Lead Time (days)",
      y = "Exceedance probability"
    )
}
```
### Overall
```{r}
cum_error |>
  prob_bust_plot()
```


### Winter
```{r}
cum_error_seasonal |>
  filter(season == "winter") |>
  prob_bust_plot()
```

### Spring
```{r}
cum_error_seasonal |>
  filter(season == "spring") |>
  prob_bust_plot()
```

### Summer
```{r}
cum_error_seasonal |>
  filter(season == "summer") |>
  prob_bust_plot()
```

### Fall
```{r}
cum_error_seasonal |>
  filter(season == "fall") |>
  prob_bust_plot()
```


```{r}
## Space remaining until upper threshold {.tabset}
### 2.5th to 97.5th percentile (Middle 95%)
# ggplot(cum_error |> filter(lead_time > 1)) +
#   geom_hline(yintercept = c(-threshold_elev, threshold_elev), linetype = "dashed") +
#   geom_errorbar(aes(lead_time / 24,
#     y = dist_to_upper_thresh_q50,
#     ymin = dist_to_upper_thresh_q05,
#     ymax = dist_to_upper_thresh_q95, color = source
#   )) +
#   # geom_line(aes(lead_time / 24, y = dist_to_nearest_thresh_q50, color = source)) +
#   geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q50, color = source)) +
#   # geom_line(aes(lead_time / 24, y = dist_to_lower_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q75, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q25, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q75, color = source)) +
#   scale_color_manual("Forecast Source", values = cols) +
#   scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
#   coord_cartesian(ylim = c(-threshold_elev, threshold_elev)) +
#   scale_y_continuous(
#     # limits = c(-threshold_elev, threshold_elev),
#     breaks = seq(-threshold_elev, threshold_elev, by = 0.1),
#     minor_breaks = NULL
#   ) +
#   facet_grid(source ~ location) +
#   theme_bw() +
#   theme(legend.position = "top") +
#   labs(
#     x = "Lead Time (days)",
#     y = "Cumulative Error (feet)"
#   )
```

```{r}
### 25th to 75th percentile (Middle 50%)
# ggplot(cum_error |> filter(lead_time > 1)) +
#   geom_hline(yintercept = c(-threshold_elev, threshold_elev), linetype = "dashed") +
#   geom_errorbar(aes(lead_time / 24,
#     y = dist_to_upper_thresh_q50,
#     ymin = dist_to_upper_thresh_q25,
#     ymax = dist_to_upper_thresh_q75,
#     color = source
#   )) +
#   # geom_line(aes(lead_time / 24, y = dist_to_nearest_thresh_q50, color = source)) +
#   geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q50, color = source)) +
#   # geom_line(aes(lead_time / 24, y = dist_to_lower_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q75, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q25, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q75, color = source)) +
#   scale_color_manual("Forecast Source", values = cols) +
#   scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
#   scale_y_continuous(
#     limits = c(-threshold_elev, threshold_elev),
#     breaks = seq(-threshold_elev, threshold_elev, by = 0.1),
#     minor_breaks = NULL
#   ) +
#   # coord_cartesian(ylim = c(-6000, 6000)) +
#   facet_grid(source ~ location) +
#   theme_bw() +
#   theme(legend.position = "top") +
#   labs(
#     x = "Lead Time (days)",
#     y = "Cumulative Error (feet)"
# )
```



```{r}
## Space remaining until lower threshold {.tabset}
### 2.5th to 97.5th percentile (Middle 95%)
# ggplot(cum_error |> filter(lead_time > 1)) +
#   geom_hline(yintercept = c(-threshold_elev, threshold_elev), linetype = "dashed") +
#   geom_errorbar(aes(lead_time / 24,
#     y = dist_to_lower_thresh_q50,
#     ymin = dist_to_lower_thresh_q05,
#     ymax = dist_to_lower_thresh_q95, color = source
#   )) +
#   # geom_line(aes(lead_time / 24, y = dist_to_nearest_thresh_q50, color = source)) +
#   geom_point(aes(lead_time / 24, y = dist_to_lower_thresh_q50, color = source)) +
#   # geom_line(aes(lead_time / 24, y = dist_to_lower_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q75, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q25, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q75, color = source)) +
#   scale_color_manual("Forecast Source", values = cols) +
#   scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
#   coord_cartesian(ylim = c(-threshold_elev, threshold_elev)) +
#   scale_y_continuous(
#     # limits = c(-threshold_elev, threshold_elev),
#     breaks = seq(-threshold_elev, threshold_elev, by = 0.1),
#     minor_breaks = NULL
#   ) +
#   facet_grid(source ~ location) +
#   theme_bw() +
#   theme(legend.position = "top") +
#   labs(
#     x = "Lead Time (days)",
#     y = "Cumulative Error (cfs)"
#   )
```


```{r}
### 25th to 75th percentile (Middle 50%)
# ggplot(cum_error |> filter(lead_time > 1)) +
#   geom_hline(yintercept = c(-threshold_elev, threshold_elev), linetype = "dashed") +
#   geom_errorbar(aes(lead_time / 24,
#     y = dist_to_lower_thresh_q50,
#     ymin = dist_to_lower_thresh_q25,
#     ymax = dist_to_lower_thresh_q75,
#     color = source
#   )) +
#   # geom_line(aes(lead_time / 24, y = dist_to_nearest_thresh_q50, color = source)) +
#   geom_point(aes(lead_time / 24, y = dist_to_lower_thresh_q50, color = source)) +
#   # geom_line(aes(lead_time / 24, y = dist_to_lower_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q50, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_upper_thresh_q75, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q25, color = source)) +
#   # geom_point(aes(lead_time / 24, y = dist_to_nearest_thresh_q75, color = source)) +
#   scale_color_manual("Forecast Source", values = cols) +
#   scale_x_continuous(limits = c(0, 3), breaks = 0:10, minor_breaks = NULL) +
#   coord_cartesian(ylim = c(-threshold_elev, threshold_elev)) +
#   scale_y_continuous(
#     # limits = c(-threshold_elev, threshold_elev),
#     breaks = seq(-threshold_elev, threshold_elev, by = 0.1),
#     minor_breaks = NULL
#   ) +
#   facet_grid(source ~ location) +
#   theme_bw() +
#   theme(legend.position = "top") +
#   labs(
#     x = "Lead Time (days)",
#     y = "Cumulative Error (cfs)"
#   )
```
