---
title: "GRH Forecast Evaluation"
subtitle: "`r params$subtitle`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: sandstone
    collapsed: true
    css: evaluation.css
params:
  # NULL for the full evaluation. A named list of cfs caps per location produces
  # the low-flow variant (see render-evaluation.R).
  flow_caps: null
  # Kept out of the YAML expression above on purpose: pandoc drops the whole
  # metadata field if the rendered subtitle contains a "<", so the caller
  # supplies plain prose.
  subtitle: "All flow conditions"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
# All evaluation points present in the data.
all_locs <- c("wilder", "bellows", "vernon", "ottauquechee", "sugar", "white")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  fig.width = 12,
  fig.height = 3.6,
  out.width = "100%"
)

library(tidyverse)
suppressMessages(import::from(
  hydroGOF,
  gof,
  NSE,
  KGE,
  me,
  mse,
  pbias,
  nrmse,
  rPearson,
  rSD,
  mNSE
))
import::from(kableExtra, kbl, kable_minimal, kable_styling)

# Palette, linetypes, shapes, location labels and theme are shared with the
# manuscript figure scripts (make-metric-figures.R, make-revision-figures.R) so
# the dashboard and the paper cannot drift apart. This file also supplies
# `locations` (the three dams the evaluation reports on -- the tributaries stay
# in the data because they feed the mainstem inflows), `product_levels` and
# `source_levels`.
source("figure-style.R")
```

Report created: `r now()`

## Metric Descriptions

- *NSE* -  Nash-Sutcliffe Efficiency, a metric that measures performance relative to the mean of the data, values range from [-∞ , 1], with values greater than 0 indicating better performance than the mean, and a perfect forecast will have a value of 1. 
- *KGE* - Kling-Gupta Efficiency, a metric that is often used similarly to NSE, but addresses several shortcomings by explicitly incorporating variability and correlation between simulations and observations. Values can range from [-∞ , 1] with negative values generally considered to indicate poor performance. 
- *PBIAS* - Percent Bias, Bias of the simulation relative to the observations, represented as a percent. Negative values indicate a low bias and positive values indicate a high bias, values closer to 0 are preferable. 
- *nRMSE* - Normalized root mean square error, expressed as a percent, lower values are better. 
- *r* - Pearson's correlation, a measure of how well the forecast and observations are correlated, values closer to 1 indicate better correlation. 
- *rSD* - Ratio of standard deviations of forecast and observed. Values close to 1 are preferable as they indicate that the forecast has the same variability as the observations. 


```{r, include=FALSE}
forecasts_0h <- read_csv(
  "processed-data/forecasts_00h.csv",
  show = F,
  progress = F
)
forecasts_6h <- read_csv(
  "processed-data/forecasts_06h.csv",
  show = F,
  progress = F
)
forecasts_12h <- read_csv(
  "processed-data/forecasts_12h.csv",
  show = F,
  progress = F
)
forecasts_24h <- read_csv(
  "processed-data/forecasts_24h.csv",
  show = F,
  progress = F
)
# The in-house GRH product is supplied as forebay targets rather than as
# forecasts, so it is renamed into the `_fc` columns and stacked with the two
# commercial products.
target <- read_csv("processed-data/targets_24h.csv", show = F, progress = F) |>
  add_column(
    source = "GRH"
  ) |>
  rename(
    wilder_fc = wilder_tgt,
    bellows_fc = bellows_tgt,
    vernon_fc = vernon_tgt,
    white_fc = white_tgt,
    ottauquechee_fc = ottauquechee_tgt,
    sugar_fc = sugar_tgt
  )
forecasts_24h <- forecasts_24h |>
  bind_rows(target)

# Optional low-flow restriction. `params$flow_caps` is NULL for the full
# evaluation; the capped variant passes a named list of cfs caps per location.
# The cap is applied per day, not per hour: if the observed flow at a location
# exceeds its cap at any hour of a day, that whole day is dropped for that
# location. Blanking `_fc` is enough because the pipelines below `na.omit()`,
# and the `perfect` and `persistence` benchmarks are derived after that, so they
# see the same restricted sample as the products.
apply_flow_caps <- function(x, caps) {
  if (!length(caps)) {
    return(x)
  }
  x <- x |> mutate(valid_date = floor_date(valid_time, "day"))
  for (loc in names(caps)) {
    days_over <- x |>
      filter(.data[[paste0(loc, "_obs")]] > caps[[loc]]) |>
      pull(valid_date) |>
      unique()
    x <- x |>
      mutate(
        "{loc}_fc" := if_else(
          valid_date %in% days_over,
          NA,
          .data[[paste0(loc, "_fc")]]
        )
      )
  }
  x |> select(-valid_date)
}

forecasts_24h <- forecasts_24h |> apply_flow_caps(params$flow_caps)

now <- now()

forecasts <- forecasts_24h |>
  mutate(
    source = case_when(
      source == "A" ~ "A",
      source == "B" ~ "B",
      source == "GRH" ~ "GRH",
      .default = NA
    )
  ) |>
  filter(
    issue_time >= ISOdate(2023, 6, 1, tz = "US/Eastern"),
    issue_time <=
      ISOdate(year(now), month(floor_date(now, "month")), 1, tz = "US/Eastern"),
  ) |>
  mutate(
    season = case_when(
      month(issue_time) %in% c(12, 1, 2) ~ "winter",
      month(issue_time) %in% 3:5 ~ "spring",
      month(issue_time) %in% 6:8 ~ "summer",
      month(issue_time) %in% 9:11 ~ "fall"
    )
  )

hourly <-
  forecasts |>
  group_by(issue_time) |>
  pivot_longer(
    -c(valid_time, issue_time, source, season),
    names_sep = "_",
    names_to = c("location", "type")
  ) |>
  mutate(location = factor(location, levels = all_locs)) |>
  mutate(
    lead_time = difftime(valid_time, issue_time, units = "hours") |>
      as.numeric()
  ) |>
  pivot_wider(
    id_cols = c(valid_time, issue_time, source, location, lead_time, season),
    names_from = type
  ) |>
  na.omit() |>
  mutate(error = fc - obs) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      mutate(
        source = "perfect",
        fc = obs,
        error = fc - obs
      )
  ) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      group_by(location, issue_time) |>
      mutate(
        source = "persistence",
        fc = obsprev[1],
        error = fc - obs
      )
  ) |>
  filter(location %in% locations) |>
  # Fix the legend order and let the shared scales key off named values.
  mutate(source = factor(source, levels = source_levels))

forecasts |>
  distinct(valid_time, source, .keep_all = T) |>
  select(-issue_time, -season) |>
  pivot_longer(
    -c(valid_time, source),
    names_sep = "_",
    names_to = c("location", "type")
  ) |>
  mutate(valid_date = floor_date(valid_time, "day")) |>
  select(-valid_time) |>
  group_by(valid_date, location, type) |>
  mutate(
    location = factor(location, levels = all_locs)
  ) |>
  group_by(valid_date, source, location, type) |>
  summarise(value = mean(value), .groups = "drop") |>
  pivot_wider(id_cols = c(valid_date, source, location), names_from = type) |>
  mutate(obsprev = lag(obs)) |>
  select(-fc, -obs) |>
  na.omit() -> obs_prev_day

daily <-
  forecasts |>
  group_by(issue_time) |>
  mutate(valid_date = floor_date(valid_time, "day")) |>
  pivot_longer(
    -c(valid_time, issue_time, source, valid_date, season),
    names_sep = "_",
    names_to = c("location", "type")
  ) |>
  mutate(
    location = factor(location, levels = all_locs)
  ) |>
  group_by(valid_date, issue_time, source, location, type, season) |>
  summarise(value = mean(value), .groups = "drop") |>
  mutate(
    lead_time = difftime(valid_date, issue_time, units = "hours") |>
      as.numeric()
  ) |>
  left_join(obs_prev_day, by = join_by(valid_date, source, location)) |>
  mutate(value = ifelse(type == "obsprev", obsprev, value)) |>
  select(-obsprev) |>
  pivot_wider(
    id_cols = c(valid_date, issue_time, source, location, lead_time, season),
    names_from = type
  ) |>
  na.omit() |>
  mutate(error = fc - obs) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      mutate(
        source = "perfect",
        fc = obs,
        error = fc - obs
      )
  ) %>%
  bind_rows(
    . |>
      filter(source == "B") |>
      group_by(location, issue_time) |>
      mutate(
        source = "persistence",
        fc = obsprev[1],
        error = fc - obs
      )
  ) |>
  filter(location %in% locations) |>
  mutate(source = factor(source, levels = source_levels))


n_day_ahead <- function(x, n) {
  x |> filter(lead_time <= n * 24, lead_time > (n - 1) * 24)
}

# GRH targets are issued a day earlier than the commercial products, so the GRH
# lead time is offset by 23 h before the day-ahead window is taken. Same
# convention as make-metric-figures.R and verify-all.R.
daily_offset_grh <- daily |>
  mutate(lead_time = ifelse(source == 'GRH', lead_time + 23, lead_time))
day1_ave_forecasts <- daily_offset_grh |> n_day_ahead(1)
day2_ave_forecasts <- daily_offset_grh |> n_day_ahead(2)
day3_ave_forecasts <- daily_offset_grh |> n_day_ahead(3)

fc_stats <- function(x) {
  x |>
    group_by(source, location, lead_time) |>
    mutate(n = n()) |>
    filter(n > 1) |>
    summarise(
      NSE = NSE(fc, obs),
      KGE = KGE(fc, obs),
      PBIAS = pbias(fc, obs),
      nRMSE = nrmse(fc, obs),
      r = rPearson(fc, obs),
      rSD = rSD(fc, obs),
      .groups = "drop"
    )
}

fc_stats_by_subgroup <- function(x, subgroup) {
  x |>
    mutate(month = month(issue_time)) |>
    group_by(source, location, lead_time, !!as.name(subgroup)) |>
    mutate(n = n()) |>
    filter(n > 1) |>
    summarise(
      NSE = NSE(fc, obs),
      KGE = KGE(fc, obs),
      PBIAS = pbias(fc, obs),
      nRMSE = nrmse(fc, obs),
      r = rPearson(fc, obs),
      rSD = rSD(fc, obs),
      .groups = "drop"
    )
}

fc_stats_daily <- daily |> fc_stats()
fc_stats_hourly <- hourly |> fc_stats()
fc_stats_da <- day1_ave_forecasts |>
  fc_stats() |>
  filter(lead_time >= 20)

fc_stats_da_season <- day1_ave_forecasts |>
  fc_stats_by_subgroup("season") |>
  filter(lead_time >= 20)

fc_stats_daily_by_month <- daily |> fc_stats_by_subgroup("month")
fc_stats_hourly_by_month <- hourly |> fc_stats_by_subgroup("month")
fc_stats_da_by_month <- day1_ave_forecasts |>
  fc_stats_by_subgroup("month") |>
  filter(lead_time >= 20)

fc_stats_daily_by_season <- daily |> fc_stats_by_subgroup("season")
fc_stats_hourly_by_season <- hourly |> fc_stats_by_subgroup("season")
```

### Day Ahead Forecast Statistics {.tabset}
#### Overall
```{r}
fc_stats_da |>
  arrange(location, source) |>
  select(-lead_time) |>
  # filter(!(source %in% c('perfect','persistence'))) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Winter 
```{r}
fc_stats_da_season |>
  filter(season == "winter") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Spring 
```{r}
fc_stats_da_season |>
  filter(season == "spring") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Summer 
```{r}
fc_stats_da_season |>
  filter(season == "summer") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

#### Fall 
```{r}
fc_stats_da_season |>
  filter(season == "fall") |>
  arrange(location, source) |>
  select(-lead_time) |>
  kbl(digits = 3) |>
  kable_styling("striped")
```

### Error Distributions (Forecast - Observed) {.tabset}
```{r}
plot_error_boxplots <- function(fc) {
  fc |>
    group_by(source, location) |>
    filter(error < quantile(error, 0.95), error > quantile(error, 0.05)) |>
    bind_rows(fc |> filter(source == "perfect")) |>
    ungroup() |>
    mutate(location = label_locations(location)) |>
    ggplot() +
    geom_hline(yintercept = 0, color = "grey35") +
    geom_boxplot(aes(source, error, fill = source)) +
    facet_wrap(vars(location), nrow = 1, scales = "free") +
    scale_fill_manual("Forecast", values = source_colors, drop = FALSE) +
    theme_paper() +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 20, hjust = 1)
    ) +
    labs(y = "Daily average flow error (cfs)", x = "Forecast")
}
```

#### Overall 
```{r}
day1_ave_forecasts |> plot_error_boxplots()
```

#### Winter 
```{r}
day1_ave_forecasts |>
  filter(season == "winter") |>
  plot_error_boxplots()
```

#### Spring 
```{r}
day1_ave_forecasts |>
  filter(season == "spring") |>
  plot_error_boxplots()
```

#### Summer 
```{r}
day1_ave_forecasts |>
  filter(season == "summer") |>
  plot_error_boxplots()
```

#### Fall 
```{r}
day1_ave_forecasts |>
  filter(season == "fall") |>
  plot_error_boxplots()
```


### Day Ahead Observed vs. Forecast Scatterplot {.tabset}
```{r}
plot_obs_vs_fc <- function(fc) {
  fc |>
    ggplot() +
    geom_abline(slope = 1, color = "grey35") +
    geom_point(aes(obs, fc, color = source), alpha = 0.4, size = 0.7) +
    facet_wrap(vars(source), nrow = 1) +
    scale_color_manual("Forecast", values = source_colors, drop = FALSE) +
    # Five panels in a row leaves no width for full cfs labels, so both axes are
    # in thousands.
    scale_x_continuous(labels = \(x) x / 1000) +
    scale_y_continuous(labels = \(x) x / 1000) +
    theme_paper() +
    theme(legend.position = "none") +
    coord_equal(xlim = c(0, 80000), ylim = c(0, 80000)) +
    labs(y = "Forecast (1000 cfs)", x = "Observed (1000 cfs)")
}
```

#### Overall 
```{r}
day1_ave_forecasts |> plot_obs_vs_fc()
```

#### Winter 
```{r}
day1_ave_forecasts |>
  filter(season == "winter") |>
  plot_obs_vs_fc()
```

#### Spring 
```{r}
day1_ave_forecasts |>
  filter(season == "spring") |>
  plot_obs_vs_fc()
```

#### Summer 
```{r}
day1_ave_forecasts |>
  filter(season == "summer") |>
  plot_obs_vs_fc()
```

#### Fall 
```{r}
day1_ave_forecasts |>
  filter(season == "fall") |>
  plot_obs_vs_fc()
```

### Day Ahead Performance by Month {.tabset}

Paper Figure 4 is this plot for the four metrics below.

```{r}
# Mirrors stat_by_month() in make-metric-figures.R (paper Figure 4).
stat_by_month <- function(x, stat) {
  x |>
    mutate(
      location = label_locations(location),
      plot_month = ISOdate(ifelse(month >= 6, 2000, 2001), month, 1)
    ) |>
    ggplot() +
    # The no-skill threshold differs by metric: using the mean of the
    # observations as a predictor gives NSE = 0 but KGE = 1 - sqrt(2) ~ -0.41
    # (Knoben et al., 2019), so the reference line must be metric-specific.
    geom_hline(
      yintercept = ifelse(stat == "KGE", -0.41, 0),
      linetype = "dashed",
      color = "grey35"
    ) +
    geom_line(
      aes(plot_month, !!as.name(stat), color = source, linetype = source),
      linewidth = 0.8
    ) +
    facet_wrap(vars(location), nrow = 1) +
    scale_x_datetime(
      date_breaks = "1 month",
      date_labels = "%b",
      minor_breaks = NULL
    ) +
    scale_source() +
    theme_paper() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
    labs(x = "", y = stat)
}
```

#### Kling-Gupta Efficiency (KGE)
```{r}
fc_stats_da_by_month |>
  stat_by_month("KGE") +
  coord_cartesian(ylim = c(-0.41, 1))
```

#### Nash-Suttecliffe Efficiency (NSE)
```{r}
fc_stats_da_by_month |>
  stat_by_month("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Pearson's Correlation (r)
```{r}
fc_stats_da_by_month |>
  stat_by_month("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Percent Bias
```{r}
fc_stats_da_by_month |>
  stat_by_month("PBIAS") +
  coord_cartesian(ylim = c(-100, 100))
```

## Daily forecast performance vs lead time {.tabset}

Paper Figure 3 is the "Overall" tab of each metric below.

```{r}
# Mirrors stat_over_lead() in make-metric-figures.R (paper Figure 3).
stat_plot_over_time <- function(x, stat) {
  x |>
    mutate(location = label_locations(location)) |>
    ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey35") +
    geom_line(
      aes(lead_time / 24, !!as.name(stat), color = source, linetype = source),
      linewidth = 0.8
    ) +
    scale_x_continuous(breaks = 0:10, minor_breaks = NULL) +
    facet_wrap(vars(location), nrow = 1) +
    scale_source() +
    theme_paper() +
    labs(x = "Lead time (days)", y = stat)
}
```

### Kling-Gupta Efficiency (KGE) {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-0.41, 1))
```

#### Winter
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Spring
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Summer
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Fall
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

### Nash-Suttecliffe Efficiency (NSE) {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Winter
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Spring
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Summer
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```

#### Fall
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-2, 1))
```


### Pearson's Correlation (r) {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Winter
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Spring
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Summer
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Fall
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```


### Percent Bias {.tabset}

#### Overall 
```{r}
fc_stats_daily |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Winter 
```{r}
fc_stats_daily_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Spring 
```{r}
fc_stats_daily_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Summer 
```{r}
fc_stats_daily_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Fall 
```{r}
fc_stats_daily_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

## Hourly forecast performance vs lead time {.tabset}

### Kling-Gupta Efficiency (KGE) {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Winter
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Spring
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Summer
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Fall
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("KGE") +
  coord_cartesian(ylim = c(-1, 1))
```
### Nash-Suttecliffe Efficiency (NSE) {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Winter
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Spring
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Summer
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

#### Fall
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("NSE") +
  coord_cartesian(ylim = c(-1, 1))
```

### Pearson's Correlation (r) {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Winter
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Spring
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Summer
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```

#### Fall
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("r") +
  coord_cartesian(ylim = c(0, 1))
```


### Percent Bias {.tabset}

#### Overall 
```{r}
fc_stats_hourly |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Winter 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "winter") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Spring 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "spring") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Summer 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "summer") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

#### Fall 
```{r}
fc_stats_hourly_by_season |>
  filter(season == "fall") |>
  stat_plot_over_time("PBIAS") +
  coord_cartesian(ylim = c(-75, 50))
```

## Cumulative Error {.tabset}

Cumulative inflow error is what moves the forebay, so it is accumulated over each
issue's own delivered horizon rather than over absolute lead time. Forecast A is
issued at 04:00 but for most of its issues the delivered series does not start
until lead hour 25, so an absolute-lead axis compares 8 accumulated hours of A
against 32 of B and produces a non-monotonic A curve driven purely by sample
composition. Indexing by position in the horizon keeps every row and makes each
point a like-for-like N-hour accumulation. This is the same construction as
paper Figures 5 and 6 (`make-revision-figures.R`).

Only the three forecast products appear here. `perfect` accumulates exactly zero
by construction and `persistence` is a day-ahead benchmark, so neither is
informative on an hourly accumulation.

```{r}
# Forebay band, in cfs-hours and in feet. 15,000 cfs-hours of inflow error moves
# the forebay by 0.5 ft at 30,000 cfs-hours per foot.
threshold <- 15000
cfs_hours_per_foot <- 30000
threshold_elev <- threshold / cfs_hours_per_foot

# GRH targets extend only 32 h past issue, so 32 h is the longest horizon over
# which all three products can be compared.
max_horizon_hour <- 32

cum_error_raw <- hourly |>
  filter(source %in% product_levels, lead_time > 0) |>
  mutate(source = factor(as.character(source), levels = product_levels)) |>
  arrange(location, source, issue_time, lead_time) |>
  group_by(location, source, issue_time) |>
  mutate(
    horizon_hour = row_number(),
    cum_error = cumsum(error)
  ) |>
  ungroup() |>
  filter(horizon_hour <= max_horizon_hour)

summarise_cum_error <- function(x, ...) {
  x |>
    group_by(location, source, horizon_hour, ...) |>
    summarise(
      n_issues = n(),
      mean_lead = mean(lead_time),
      ave_cum_error = mean(cum_error),
      q95_cum_error = quantile(cum_error, 0.975),
      q75_cum_error = quantile(cum_error, 0.75),
      q50_cum_error = quantile(cum_error, 0.50),
      q25_cum_error = quantile(cum_error, 0.25),
      q05_cum_error = quantile(cum_error, 0.025),
      p_bust = mean(abs(cum_error) > threshold),
      .groups = "drop"
    ) |>
    mutate(location = label_locations(location))
}

cum_error <- cum_error_raw |> summarise_cum_error()
cum_error_seasonal <- cum_error_raw |> summarise_cum_error(season)
```

### 2.5th to 97.5th percentile (Middle 95%) {.tabset}

Paper Figure 5 is the "Overall" tab.

```{r}
# Mirrors figure_5 in make-revision-figures.R. `lower`/`upper` name the band
# columns so the 95% and 50% versions share one function.
cum_error_plot <- function(
  error,
  lower = "q05_cum_error",
  upper = "q95_cum_error",
  band = threshold_elev,
  scale = cfs_hours_per_foot
) {
  ggplot(error, aes(x = horizon_hour, color = source, fill = source)) +
    geom_hline(
      yintercept = c(-band, band),
      linetype = "dashed",
      color = "grey35",
      linewidth = 0.4
    ) +
    # Outline each band and drop the fill to 10%. The outline is what makes the
    # bands readable where they overlap: three faint fills with no edge blend
    # into intermediate colors that are in no legend entry.
    geom_ribbon(
      aes(
        ymin = !!as.name(lower) / scale,
        ymax = !!as.name(upper) / scale,
        linetype = source
      ),
      alpha = 0.10,
      linewidth = 0.45
    ) +
    geom_line(aes(y = q50_cum_error / scale), linewidth = 0.9) +
    facet_wrap(vars(location), nrow = 1) +
    scale_source(fill = TRUE) +
    scale_x_continuous(
      breaks = seq(8, max_horizon_hour, 8),
      limits = c(1, max_horizon_hour),
      expand = expansion(mult = 0.02)
    ) +
    theme_paper() +
    labs(
      x = "Hours into forecast horizon",
      y = "Cumulative forebay error (ft)"
    )
}

cum_error_plot50 <- function(error) {
  cum_error_plot(error, lower = "q25_cum_error", upper = "q75_cum_error")
}
```

#### Overall
```{r, fig.height=4.6}
cum_error |>
  cum_error_plot()
```


#### Winter
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "winter") |>
  cum_error_plot()
```

#### Spring
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "spring") |>
  cum_error_plot()
```

#### Summer
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "summer") |>
  cum_error_plot()
```

#### Fall
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "fall") |>
  cum_error_plot()
```

### 25th to 75th percentile (Middle 50%) {.tabset}

The same plot over the interquartile range instead of the middle 95%.

#### Overall
```{r, fig.height=4.6}
cum_error |>
  cum_error_plot50()
```


#### Winter
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "winter") |>
  cum_error_plot50()
```

#### Spring
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "spring") |>
  cum_error_plot50()
```

#### Summer
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "summer") |>
  cum_error_plot50()
```

#### Fall
```{r, fig.height=4.6}
cum_error_seasonal |>
  filter(season == "fall") |>
  cum_error_plot50()
```

## Probability of exceeding 0.5 ft cumulative error {.tabset}

Paper Figure 6 is the "Overall" tab, restricted to the marked hours.

```{r}
# Mirrors figure_6 in make-revision-figures.R. The paper plots the four marker
# hours only; the dashboard keeps the whole curve and marks those hours.
report_hours <- c(8, 16, 24, 32)

prob_bust_plot <- function(error) {
  ggplot(error, aes(x = horizon_hour, y = p_bust, color = source)) +
    geom_line(aes(linetype = source), linewidth = 0.9) +
    geom_point(
      data = function(d) filter(d, horizon_hour %in% report_hours),
      aes(shape = source),
      size = 2.4
    ) +
    facet_wrap(vars(location), nrow = 1) +
    scale_source(shape = TRUE) +
    scale_x_continuous(
      breaks = report_hours,
      limits = c(1, max_horizon_hour),
      expand = expansion(mult = 0.04)
    ) +
    scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1)) +
    theme_paper() +
    labs(
      x = "Hours into forecast horizon",
      y = "Probability outside forebay band"
    )
}
```
### Overall
```{r}
cum_error |>
  prob_bust_plot()
```


### Winter
```{r}
cum_error_seasonal |>
  filter(season == "winter") |>
  prob_bust_plot()
```

### Spring
```{r}
cum_error_seasonal |>
  filter(season == "spring") |>
  prob_bust_plot()
```

### Summer
```{r}
cum_error_seasonal |>
  filter(season == "summer") |>
  prob_bust_plot()
```

### Fall
```{r}
cum_error_seasonal |>
  filter(season == "fall") |>
  prob_bust_plot()
```
